### 2016-11-30 Paper_v1.tex
Nov30_2

% Begin ...
\begin{document}
\begin{titlepage}
{\par\centering\textbf{\Large {Toward a Genome Scale Dynamic Model of Cell-Free Protein Synthesis in \emph{Escherichia~coli}}}}
\vspace{0.05in}
{\par \centering \large{Nicholas Horvath, Michael Vilkhovoy, Joseph Wayman, Kara Calhoun$^{1}$, James Swartz$^{1}$ and Jeffrey D. Varner$^{*}$}}
\vspace{0.10in}
{\par \centering {Robert Frederick Smith School of Chemical and Biomolecular Engineering}}
{\par \centering {Cornell University, Ithaca NY 14853}}
\vspace{0.1in}
{\par \centering {$^{1}$School of Chemical Engineering}}
{\par \centering {Stanford University, Stanford, CA 94305}}
{\par \centering \textbf{Running Title:}~Dynamic modeling of cell-free protein synthesis}
\vspace{0.1in}
{\par \centering \textbf{To be submitted:}~\emph{Scientific~Reports}}
\vspace{0.5in}
{\par \centering $^{*}$Corresponding author:}
{\par \centering Jeffrey D. Varner,}
{\par \centering Professor, Robert Frederick Smith School of Chemical and Biomolecular Engineering,}
{\par \centering 244 Olin Hall, Cornell University, Ithaca NY, 14853}
{\par \centering Email: jdv27@cornell.edu}
{\par \centering Phone: (607) 255 - 4258}
{\par \centering Fax: (607) 255 - 9166}
\end{titlepage}
\date{}
\thispagestyle{empty}
\pagebreak
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Abstract}
Cell-free protein expression systems have become widely used in systems and synthetic biology.
In this study, we developed an ensemble of dynamic \textit{E. coli} cell-free protein synthesis (CFPS) models.
Model parameters were estimated from measurements of glucose, organic acids, energy species, amino acids, and the protein product, chloramphenicol acetyltransferase (CAT).
The ensemble described all of the training data, especially the central carbon metabolism.
The model predicted a carbon yield for CAT production that was equal to 23\% of the maximum theoretical yield, calculated using sequence-specific flux balance analysis.
This suggests that CAT production could be further optimized.
The dynamic modeling approach predicted that substrate consumption of glucose and pyruvate and oxidative phosphorylation were most important to both CAT production and the system as a whole, while CAT production alone depended heavily on the CAT synthesis reaction.
Conversely, CAT production was robust to allosteric control, as was most of the network, with the exception of the organic acids in central carbon metabolism.
This study is the first to model dynamic protein production in \textit{E. coli}, and should provide a foundation for genome-scale,
dynamic modeling of cell-free \textit{E. coli} protein synthesis.

>>>>

% Begin ...
\begin{document}
\begin{titlepage}
{\par\centering\textbf{\Large {Toward a Genome Scale Dynamic Model of Cell-Free Protein Synthesis in \emph{Escherichia~coli}}}}
\vspace{0.05in}
{\par \centering \large{Nicholas Horvath, Michael Vilkhovoy, Joseph Wayman, Kara Calhoun$^{1}$, James Swartz$^{1}$ and Jeffrey D. Varner$^{*}$}}
\vspace{0.10in}
{\par \centering {Robert Frederick Smith School of Chemical and Biomolecular Engineering}}
{\par \centering {Cornell University, Ithaca NY 14853}}
\vspace{0.1in}
{\par \centering {$^{1}$School of Chemical Engineering}}
{\par \centering {Stanford University, Stanford, CA 94305}}
{\par \centering \textbf{Running Title:}~Dynamic modeling of cell-free protein synthesis}
\vspace{0.1in}
{\par \centering \textbf{To be submitted:}~\emph{Scientific~Reports}}
\vspace{0.5in}
{\par \centering $^{*}$Corresponding author:}
{\par \centering Jeffrey D. Varner,}
{\par \centering Professor, Robert Frederick Smith School of Chemical and Biomolecular Engineering,}
{\par \centering 244 Olin Hall, Cornell University, Ithaca NY, 14853}
{\par \centering Email: jdv27@cornell.edu}
{\par \centering Phone: (607) 255 - 4258}
{\par \centering Fax: (607) 255 - 9166}
\end{titlepage}
\date{}
\thispagestyle{empty}
\pagebreak
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Abstract}
Cell-free protein expression systems have become widely used in systems and synthetic biology.
In this study, we developed an ensemble of dynamic \textit{E. coli} cell-free protein synthesis (CFPS) models.
Model parameters were estimated from measurements of glucose, organic acids, energy species, amino acids, and the protein product, chloramphenicol acetyltransferase (CAT).
The ensemble described all of the training data, especially the central carbon metabolism.
The model predicted a carbon yield for CAT production that was equal to 23\% of the maximum theoretical yield, calculated using sequence-specific flux balance analysis.
This suggests that CAT production could be further optimized.
The dynamic modeling approach predicted that substrate consumption and oxidative phosphorylation were most important to both CAT production and the system as a whole, while CAT production alone depended heavily on the CAT synthesis reaction.
Conversely, CAT production was robust to allosteric control, as was most of the network, with the exception of the organic acids in central carbon metabolism.
This study is the first to model dynamic protein production in \textit{E. coli}, and should provide a foundation for genome-scale,
dynamic modeling of cell-free \textit{E. coli} protein synthesis.
